---
title: "Validation Study of the Use of PF Ratio for Patients with Acute Hypoxaemic Respiratory Failure Admitted to Australian and New Zealand Intensive Care Units" 
subtitle: "Statistical Analysis"
author: "Dr Benjamin Moran, MBBS, BMedSci (Hons), MMedStats, FCICM, FANZCA"
date: today
format:   
  html:
    embed-resources: true
    grid: 
      body-width: 4000px
      sidebar-width: 150px
execute:
  cache: true
  echo: false
toc: true
toc-location: left
toc-depth: 5
number-sections: true
bibliography: references.bib
theme: yeti
---

## Introduction

This is an explanation of the statistical analysis for the study validating the use of PaO~2~:FiO~2~ ratio in patients admitted to ICU with acute hypoxaemic respiratory failure.

## Methods

This is a retrospective study using data from the Australian and New Zealand Intensive Care Society (ANZICS) adult ICU patient database (APD). This manuscript has been prepared and reported in accordance with the Strengthening the Reporting of Observational Studies in Epidemiology (STROBE) Statement.

### Patient Population

We included all patients in the ANZICS APD from 01/01/2017 to 31/12/2022.

### Aims and Objectives

1.  Determine the association between PaO~2~:FiO~2~ ratio and hospital mortality.
2.  Determine the association between PaO~2~:FiO~2~ ratio and ICU mortality.
3.  Determine the association between PaO~2~:FiO~2~ ratio and mortality at other timepoints (28 day, 6 months, 12 months).
4.  Validate the PaO~2~:FiO~2~ ratio as a diagnostic tool for prediction of hospital mortality, and mortality at other times (ICU, 28-day, 6 months, 12 months).
5.  Determine the validity of the PaO~2~:FiO~2~ ratio in predicting hospital mortality in pre-specified subgroups (invasive ventilation, ventilatory support, age, sex, admission diagnoses, frailty categories, presence of treatment limitations).

### Statistical Analysis

We summarized baseline ICU and patient-level characteristics and unadjusted outcomes using standard descriptive statistics. For categorical data, we used counts and percentages, and for continuous data we used mean ± standard deviation (SD) or median (interquartile range, IQR) as appropriate depending on the distribution of data.

#### Association Between PaO~2~:FiO~2~ Ratio and Hospital Mortality

The unadjusted association between PaO~2~:FiO~2~ ratio and hospital mortality was evaluated visually as a continuous, non-linear variable using restricted cubic splines with 4 knots. Specific unadjusted hospital mortality values for standard PaO~2~:FiO~2~ ratio values were also calculated through interrogation of the spline curve.

#### Validation of PaO~2~:FiO~2~ ratio in Predicting Hospital Mortality

To determine the validitiy of the PaO~2~:FiO~2~ ratio, the area under the receiver operator characteristic curve (AUCROC) was calculated. 95% confidence intervals around the AUC were calculated using 1,000 bootstrap samples.

To determine the optimal cut-off of PaO~2~:FiO~2~ ratio in predicting hospital mortality, receiver operator characteristic curves (ROC) were generated. A cut-off level representing the highest sum of sensitivity and specificity hased on each patients PaO~2~:FiO~2~ ratio was calculated using the Youden method [@youden1950]. In this method, the sensitivity and specificity was calculated over a range of PaO~2~:FiO~2~ ratios. For each value, the Youden's J index was calculated by using the following formula (Youden = Sensitivity + Specificity - 1). The value that corresponds to the highest Youden index was identified as the optimal cut-off, reflecting the highest sum of sensitivity and specifity. We also calculated the sensitivity, specificity, negative predictive value and positive predictive value for this cut-off.

#### Subgroup Analysis

Patients were analysed for validation of the PaO~2~:FiO~2~ ratio in prediction of hospital mortality in the following subgroups: receiving invasive ventilation during the index ICU admission, receiving of other respiratory support (eg non-invasive ventilation, extracorporeal membrane oxygenation) during the ICU admission, sex, age categories, admission diagnoses (medical, cardiac surgery, neurosurgery/trauma, sepsis, post-operative), frailty category (fit/well, mild, moderate, severe frailty) and the presence of treatment limitation status on ICU admission. The same methodology as above was employed.

As there were >800,000 patients in the dataset, a 2-sided p-value of 0.001 was used for statistical significance. Given that there is an increased risk of Type-1 error with multiple testing, the results of the secondary objectives should be viewed as exploratory. Hence, no adjustment for multiplicity was used. Only patients with complete data for all covariates were included in the analysis. Statistical analyses were performed using R Version 4.3.1 (R Core Team, R Foundation for Statistical Computing, Vienna, Austria) and RStudio Version 2023.12.1 (Posit Software, PBC, Boston, MA). Packages used for analysis included tidytable [@fairbanks], tidyverse [@wickham], data.table [@barrett], gtsummary [@sjoberg2021], gt [@iannone2022], cutpointr [@thiele] and pROC [@robin].

```{r setup, include=FALSE, warning=FALSE, cache.lazy=FALSE}
library(tidytable)
library(gtsummary)
library(gt)
library(cutpointr)
library(pROC)

knitr::opts_chunk$set(echo = F, message = F, warning = F, error = T, 
                      out.width = "90%", dev = "png", dpi = 300, cache = T)

# Global Options
options(mc.cores = parallel::detectCores())

# Import and Filter Dataset
ahrf_validation <- fread("~/Documents/Research/Projects/Current Projects/ANZICS APD/APD ARF Study/Data/ahrfvalidation8Sep.csv",
             select = c("DSITEID", "HospitalClassification", "PublicPrivate", "Country", "PatientID", "IcuAdmitYYYY", "ICU_AD_DTM", "SEX", "AGE", "ICU_SRCE", "AP3DIAG", "Apache3Score", "Apache2Score",  "ANZRODRiskOfDeath", "FRAILTY", "PAO2_ANZ", "FIO2_ANZ", "DIED_HOSP", "DIED_ICU", "ICU_HRS", "HOSP_HRS", "HOSP_OUTCM", "INV_IND", "INV_DAYONE", "ECMO_IND", "NIV_IND", "CHR_RESP", "CHR_CVS", "COVID_pneumonitis_derived", "TRACHE_IND", "CHR_LIV", "CHR_REN", "INV_HOURS", "HOSP_AD_DTM", "HOSP_DS_DTM", "TREAT_LMT", "EMG_RSP_ADM", "PRE_ICU_HRS", "PANDEMIC", "sofa_score_derived", "NDI_NHI_DeathDate"))



#==========================================DEMOGRAPHICS==========================================#

# Rename & Reorder Gender
ahrf_validation <- ahrf_validation |> 
  mutate(SEX = case_when(
    SEX == "F" ~ "Female",
    SEX == "M" ~ "Male",
    SEX == "I" ~ "Intersex/Indeterminate",
    SEX == "NULL" ~ "Unknown")) |> 
  mutate(SEX = factor(SEX, levels = c("Female", "Male", "Intersex/Indeterminate", "Unknown")))

# Age Category
ahrf_validation <- ahrf_validation |> mutate(age_cat = case_when(
  AGE <= 44 ~ "<44",
  AGE >44 & AGE <=64 ~ "45-64",
  AGE >64 & AGE < 84 ~ "65-84",
  AGE > 84 ~ ">84")) 

# Rename ICU Admission Source #
ahrf_validation <- ahrf_validation |>  mutate(icu_admit_source = case_when(
            (ICU_SRCE ==1) ~ "Operating Theatre/Recovery",
            (ICU_SRCE ==2) ~ "Emergency Department",
            (ICU_SRCE ==3) ~ "Ward",
            (ICU_SRCE ==4) ~ "ICU, Same Hospital",
            (ICU_SRCE ==5) ~ "Other Hospital",
            (ICU_SRCE ==6) ~ "Other Hopsital",
            (ICU_SRCE ==9) ~ "Direct from Home")) |> 
  select(-ICU_SRCE)

# Reorder Admission Source
ahrf_validation <- ahrf_validation |>  mutate(icu_admit_source = factor(icu_admit_source, levels = c("Emergency Department", "Operating Theatre/Recovery", "Ward", "ICU, Same Hospital", "Other Hospital", "Direct from Home")))

# Relabel Admission Diagnoses #
# Minimal Admission Diagnosis
ahrf_validation <- ahrf_validation |> 
  mutate(ap3diag_min = case_when(
            (AP3DIAG>=1202) & (AP3DIAG<=1213) ~ "Cardiac Surgery",
            (AP3DIAG>=1301) & (AP3DIAG<=1304) ~ "Post-Operative",
            (AP3DIAG>=1401) & (AP3DIAG<=1413) ~ "Post-Operative",
            (AP3DIAG>=1501) & (AP3DIAG<=1506) ~ "Trauma/Neurosurgery",
            (AP3DIAG>=1602) & (AP3DIAG<=1605) ~ "Trauma/Neurosurgery",
            (AP3DIAG>=1701) & (AP3DIAG<=1705) ~ "Post-Operative",
            (AP3DIAG>=1801) & (AP3DIAG<=1803) ~ "Post-Operative",
            (AP3DIAG>=1902) & (AP3DIAG<=1904) ~ "Post-Operative",
            (AP3DIAG==2101) ~ "Post-Operative",
            (AP3DIAG==2201) ~ "Post-Operative",
            (AP3DIAG>=101) & (AP3DIAG<=111) ~ "Medical",
            (AP3DIAG>=201) & (AP3DIAG<=213) ~ "Medical",
            (AP3DIAG>=301) & (AP3DIAG<=313) ~ "Medical",
            (AP3DIAG>=401) & (AP3DIAG<410) ~ "Medical",
            (AP3DIAG>=501) & (AP3DIAG<=504) ~ "Sepsis",
            (AP3DIAG>601) & (AP3DIAG<=605) ~ "Trauma/Neurosurgery",
            (AP3DIAG>=701) & (AP3DIAG<=704) ~ "Medical",
            (AP3DIAG>=801) & (AP3DIAG<=802) ~ "Medical",
            (AP3DIAG>=901) & (AP3DIAG<=903) ~ "Medical",
            (AP3DIAG>=1101) & (AP3DIAG<=1102) ~ "Medical")) 

# Reorder Apache3
ahrf_validation <- ahrf_validation |>  mutate(ap3diag_min = factor(ap3diag_min, levels = c("Medical", "Post-Operative", "Sepsis", "Trauma/Neurosurgery", "Cardiac Surgery")))

# Convert Apache3Score to numeric
ahrf_validation <- ahrf_validation |> mutate(Apache3Score = as.numeric(Apache3Score))

# Convert ANZROD to numeric
ahrf_validation <- ahrf_validation |> mutate(ANZRODRiskOfDeath = as.numeric(ANZRODRiskOfDeath))

# Reorder Hospital Type
ahrf_validation <- ahrf_validation |>  mutate(HospitalClassification = factor(HospitalClassification, levels = c("Tertiary", "Metropolitan", "Rural / Regional", "Private")))

# Categorise CFS
ahrf_validation <- ahrf_validation |> 
  mutate(FRAILTY = case_when(
    (FRAILTY == "1"|FRAILTY == "2"|FRAILTY == "3") ~ "Fit/Well",
    (FRAILTY == "4"|FRAILTY == "5") ~ "Mild",
    FRAILTY == "6" ~ "Moderate",
     FRAILTY == "7" ~ "Severe"))

#==========================================EXPOSURE==========================================#

# Calculate Continuous PaO2:FiO2
ahrf_validation <- ahrf_validation |> 
  mutate(pf_cont = as.numeric(PAO2_ANZ)/as.numeric(FIO2_ANZ))

# Calculate AHRF Vs No-AHRF
ahrf_validation <- ahrf_validation |> 
  mutate(AHRF = case_when(
    pf_cont >=300 ~ "No AHRF",
    pf_cont <300 ~ "AHRF"))

# Reorder AHRF
ahrf_validation <- ahrf_validation |> 
  mutate(AHRF = factor(AHRF, levels = c("No AHRF", "AHRF")))

#Calculate Categorical PaO2:FiO2
ahrf_validation <- ahrf_validation |> 
  mutate(pf_cat = case_when(
    pf_cont < 100 ~ "Severe",
    pf_cont >=100 & pf_cont < 200 ~ "Moderate",
    pf_cont >=200 & pf_cont <300 ~ "Mild",
    pf_cont >=300 ~ "None"))

# Reorder pf_cat
ahrf_validation <- ahrf_validation |> 
  mutate(pf_cat = factor(pf_cat, levels = c("None", "Mild", "Moderate", "Severe")))

# Rename AHRF Causes (APache3 Diagnostic Code)
ahrf_validation <- ahrf_validation |> 
  mutate(ahrf_adm = case_when(
    (AP3DIAG>=1202) & (AP3DIAG<=1213) ~ "Cardiac Surgery",
    (AP3DIAG>=1301) & (AP3DIAG<=1304) ~ "Other Post-Operative",
    (AP3DIAG>=1401) & (AP3DIAG<=1413) ~ "Gastrointestinal Surgery",
    (AP3DIAG>=1501) & (AP3DIAG<=1506) ~ "Other Post-Operative",
    (AP3DIAG>=1602) & (AP3DIAG<=1605) ~ "Other Post-Operative",
    (AP3DIAG>=1701) & (AP3DIAG<=1705) ~ "Other Post-Operative",
    (AP3DIAG>=1801) & (AP3DIAG<=1803) ~ "Other Post-Operative",
    (AP3DIAG>=1902) & (AP3DIAG<=1904) ~ "Other Post-Operative",
    (AP3DIAG==2101) ~ "Other Non-Operative",
    (AP3DIAG==2201) ~ "Other Non-Operative",
    (AP3DIAG>=101) & (AP3DIAG<=111) ~ "Cardiovascular Disease",
    (AP3DIAG>=201) & (AP3DIAG<=213) ~ "Respiratory Disease",
    (AP3DIAG>=301) & (AP3DIAG<=313) ~ "Other Non-Operative",
    (AP3DIAG>=401) & (AP3DIAG<410) ~ "Other Non-Operative",
    (AP3DIAG>=501) & (AP3DIAG<=504) ~ "Sepsis",
    (AP3DIAG>601) & (AP3DIAG<=605) ~ "Other Non-Operative",
    (AP3DIAG>=701) & (AP3DIAG<=704) ~ "Other Non-Operative",
    (AP3DIAG>=801) & (AP3DIAG<=802) ~ "Other Non-Operative",
    (AP3DIAG>=901) & (AP3DIAG<=903) ~ "Other Non-Operative",
    (AP3DIAG>=1101) & (AP3DIAG<=1102) ~ "Other Non-Operative")) 

# Refactor ahrf_cause
ahrf_validation <- ahrf_validation |> 
  mutate(ahrf_adm = factor(ahrf_adm, levels = c("Cardiac Surgery", "Respiratory Disease", "Cardiovascular Disease", "Sepsis", "Gastrointestinal Surgery", "Other Non-Operative", "Other Post-Operative")))

#==========================================OUTCOMES==========================================#

# Change LOS to days (from hours)
ahrf_validation <- ahrf_validation |>  mutate(icu_los = as.numeric(ICU_HRS)/24) |> select(-ICU_HRS)
ahrf_validation <- ahrf_validation |>  mutate(hosp_los = as.numeric(HOSP_HRS)/24) |> select(-HOSP_HRS)

library(tidyverse)
# Mortality Date Data- Create Survival Time Interval
ahrf_validation <- ahrf_validation |> 
  # Convert character to date format
  mutate(ICU_AD_DTM = ymd_hms(ICU_AD_DTM),
         NDI_NHI_DeathDate = ymd_hms(NDI_NHI_DeathDate)) |> 
  # Create interval between ICU admission and death
  mutate(surv_interval = difftime(NDI_NHI_DeathDate, ICU_AD_DTM, units = "days"))

# Create Binary variables of survival at different time-points
# 28 Day Mortality
ahrf_validation<- ahrf_validation |> 
  mutate(mort_28d = if_else(surv_interval<28, 1, 0),
         mort_90d = if_else(surv_interval<90, 1, 0),
         mort_180d = if_else(surv_interval<180, 1, 0),
         mort_365d = if_else(surv_interval<365, 1, 0))

# Convert NAs into 0
ahrf_validation<- ahrf_validation |> 
  mutate(mort_28d = replace_na(mort_28d, 0),
         mort_90d = replace_na(mort_90d, 0),
         mort_180d = replace_na(mort_180d, 0),
         mort_365d = replace_na(mort_365d, 0))


#==========================================REMOVE MISSING DATA==========================================#
# Remove variables that have missing data
ahrf_validation <- ahrf_validation |> 
  filter(!(DIED_HOSP == "NA"))

ahrf_validation <- ahrf_validation[complete.cases(ahrf_validation$pf_cont),]

# Filtering patients with PF > 700
ahrf_validation <- ahrf_validation |> 
  filter(!(pf_cont > 700))

#==========================================TABLE 1==========================================#
# Count row numbers to get patient numbers in table 1
ahrf_validation <- ahrf_validation |>  mutate(PatientID = row_number()) |>  
  group_by(PatientID) |> 
  mutate(count_pt = row_number() == 1L) |> 
  ungroup()

# Generate Table 1- AHRF Categories
t1 <- ahrf_validation |>  
  select(patientn = count_pt, AGE, age_cat, SEX, Apache2Score, Apache3Score, ANZRODRiskOfDeath, sofa_score_derived, ap3diag_min, COVID_pneumonitis_derived, icu_admit_source, HospitalClassification, CHR_RESP, CHR_CVS, CHR_LIV, CHR_REN, FRAILTY, pf_cat) |>  
    tbl_summary(by = pf_cat, missing = "no",
            label = list(
              patientn = "Number of Patients",
              AGE = "Median Age in Years (IQR)",
              age_cat = "Age Category, Years",
              SEX = "Gender",
              Apache2Score = "Median APACHE II Score (IQR)",
              Apache3Score = "Median APACHE III Score (IQR)",
              ANZRODRiskOfDeath = "Median ANZROD (IQR)",
              sofa_score_derived = "Median SOFA (IQR)",
              ap3diag_min = "Admission Diagnosis",
              COVID_pneumonitis_derived = "COVID Penumonitis (Proven)",
              icu_admit_source = "Admission Source",
              HospitalClassification = "Hospital Type",
              CHR_RESP = "Chronic Respiratory Disease",
              CHR_CVS = "Chronic CVS Disease",
              CHR_LIV = "Chronic Hepatic Disease",
              CHR_REN = "Chronic Renal Disease",
              FRAILTY = "Frailty"),
            value = list(
              COVID_pneumonitis_derived ~ 1,
              CHR_RESP ~ 1,
              CHR_CVS ~ 1,
              CHR_LIV ~ 1,
              CHR_REN ~ 1),
            statistic = list(
              patientn ~ "{n}",
              AGE ~ "{median} ({p25}-{p75})",
              SEX ~ "{n} ({p}%)",
              Apache2Score ~ "{median} ({p25}-{p75})",
              Apache3Score ~ "{median} ({p25}-{p75})",
              ANZRODRiskOfDeath ~ "{median} ({p25}-{p75})",
              sofa_score_derived ~ "{median} ({p25}-{p75})",
              ap3diag_min ~ "{n} ({p}%)",
              icu_admit_source ~ "{n} ({p}%)")) |> 
  modify_header(stat_1 = "**None (PF >300)**", stat_2 = "**Mild (PF 200-300)**", stat_3 = "**Moderate (PF 100-200)**", stat_4 = "**Severe (PF < 100)**") |> 
  modify_spanning_header(all_stat_cols() ~ "**Acute Hypoxaemic Respiratory Failure Category**") |> 
  add_overall(col_label = "**Overall**", last = FALSE) |> 
  bold_labels() |> 
  italicize_labels()
```

## Results

Of  853,598 patients admitted to 211 ICUs during the study period, 440,791 (51.6%) patients had acute hypoaxemic respiratory failure. Of this cohort, 234,003 (27.4%) had mild AHRF, 169,067 (19.8%) had moderate AHRF and 37,721 (4.4%) had severe AHRF.

### Patient Demographics

Below are the demographic tables. This table has the 4 AHRF categories (none, mild, moderate, severe) to look at the breakdown of patients within each AHRF category.

```{r, warning=FALSE, message=FALSE}
t1
```

### ICU & Hospital Outcomes
```{r, message=FALSE, warning=FALSE}
# Generate Table of all outcomes- AHRF Categories
tbl_icu_hosp_outcomes <- ahrf_validation |> 
  mutate(icu_los_med = icu_los,
         icu_los_mean = icu_los,
         hosp_los_med = hosp_los,
         hosp_los_mean = hosp_los) |> 
  select(DIED_HOSP, DIED_ICU, icu_los_mean, icu_los_med, hosp_los_mean, hosp_los_med, pf_cat) |>  
    tbl_summary(by = pf_cat, missing = "no",
            label = list(
              DIED_HOSP = "Hospital Mortality",
              DIED_ICU = "ICU Mortality",
              icu_los_mean = "Mean ICU Length of Stay in Days (SD)",
              hosp_los_mean = "Mean Hospital Length of Stay in Days (SD)",
              icu_los_med = "Median ICU Length of Stay in Days (IQR)",
              hosp_los_med = "Median Hospital Length of Stay in Days (IQR)"),
            value = list(
              DIED_HOSP ~ "1",
              DIED_ICU ~ "1"),
            statistic = list(
              DIED_HOSP ~ "{n} ({p}%)",
              DIED_ICU ~ "{n} ({p}%)",
              icu_los_mean ~ "{mean} ({sd})",
              hosp_los_mean ~ "{mean} ({sd})",
              icu_los_med ~ "{median} ({p25}-{p75})",
              hosp_los_med ~ "{median} ({p25}-{p75})")) |> 
  modify_header(stat_1 = "**None (PF >300)**", stat_2 = "**Mild (PF 200-300)**", stat_3 = "**Moderate (PF 100-200)**", stat_4 = "**Severe (PF < 100)**") |> 
  modify_spanning_header(all_stat_cols() ~ "**Acute Respiratory Failure Category**") |>
  bold_labels() |> 
  add_overall(col_label = "**Overall**", last = FALSE)

tbl_icu_hosp_outcomes
```

### Mortality Outcomes

Below are the incidences of hospital mortality, ICU mortality, 6-month mortality and 12-month mortality.

```{r, cache.lazy=FALSE, message=FALSE, warnings=FALSE}
surv_tbl <- ahrf_validation |> 
    select(patientn = count_pt, DIED_ICU, DIED_HOSP, mort_28d, mort_90d, mort_180d, mort_365d, pf_cat) |> 
  tbl_summary(by = pf_cat, missing = "no",
            label = list(
              patientn = "Number of Patients",
              DIED_ICU = "ICU Mortality",
              DIED_HOSP = "Hospital Mortality",
              mort_28d = "28-Day Mortality",
              mort_90d = "90-Day Mortality",
              mort_180d = "180-Day Mortality",
              mort_365d = "1-Year Mortality"),
            value = list(
              DIED_ICU ~ 1,
              DIED_HOSP ~ 1,
              mort_28d ~ 1,
              mort_90d ~ 1,
              mort_180d ~ 1,
              mort_365d ~ 1),
            statistic = list(
              patientn ~ "{n}"))|> 
  modify_header(stat_1 = "**None (PF >300)**", stat_2 = "**Mild (PF 200-300)**", stat_3 = "**Moderate (PF 100-200)**", stat_4 = "**Severe (PF < 100)**") |> 
  modify_spanning_header(all_stat_cols() ~ "**Acute Hypoxaemic Respiratory Failure Category**") |> 
  add_overall(col_label = "**Overall**", last = FALSE) |> 
  bold_labels() |> 
  italicize_labels()

surv_tbl
```

### Primary Outcome: Hospital Mortality

```{r, cache.lazy=FALSE, include=FALSE, warnings=FALSE}

# Plot Continuous PF and Unadjusted Mortality
mort_pf_curve <- ggplot(data=ahrf_validation, aes(x=pf_cont, y=DIED_HOSP))+
  geom_smooth(data = ahrf_validation, method = "glm", formula = y ~ splines::ns(x, 4), se=TRUE)+
  labs(x=expression(paste(PaO[2]:FiO[2], " Ratio")), y="Hospital Mortality (%)", title="Unadjusted Association of Day 1 PF Ratio and Hospital Mortality")+
  coord_cartesian(ylim = c(0,0.45))+
  scale_y_continuous(labels = scales::percent)+
  theme_bw() +
  theme(panel.grid.major.x = element_blank())

# Access smoothed fitted values
smooths <- ggplot_build(mort_pf_curve)$data[[1]]

# Find values for varying pf values
x_50 <- approx(x = smooths$x, y = smooths$y, xout = 50)
x_100 <- approx(x = smooths$x, y = smooths$y, xout = 100)
x_150 <- approx(x = smooths$x, y = smooths$y, xout = 150)
x_200 <- approx(x = smooths$x, y = smooths$y, xout = 200)
x_300 <- approx(x = smooths$x, y = smooths$y, xout = 300)
x_400 <- approx(x = smooths$x, y = smooths$y, xout = 400)
x_700 <- approx(x = smooths$x, y = smooths$y, xout = 700)

# Plot with lines
mort_pf_values <-
  ggplot(data=ahrf_validation, aes(x=pf_cont, y=DIED_HOSP))+
  geom_smooth(data = ahrf_validation, method = "glm", formula = y ~ splines::ns(x, 4), se=TRUE)+
  geom_segment(x=x_100$x, xend=x_100$x, y=-Inf, yend=x_100$y, colour="#2299FF")+
  geom_segment(x=-Inf, xend=x_100$x, y=x_100$y, yend=x_100$y, colour="#2299FF")+
  geom_segment(x=x_200$x, xend=x_200$x, y=-Inf, yend=x_200$y, colour="#2299FF")+
  geom_segment(x=-Inf, xend=x_200$x, y=x_200$y, yend=x_200$y, colour="#2299FF")+
  geom_segment(x=x_300$x, xend=x_300$x, y=-Inf, yend=x_300$y, colour="#2299FF")+
  geom_segment(x=-Inf, xend=x_300$x, y=x_300$y, yend=x_300$y, colour="#2299FF")+
  geom_segment(x=x_400$x, xend=x_400$x, y=-Inf, yend=x_400$y, colour="#2299FF")+
  geom_segment(x=x_700$x, xend=x_700$x, y=-Inf, yend=x_700$y, colour="#2299FF")+
  geom_segment(x=-Inf, xend=x_700$x, y=x_700$y, yend=x_700$y, colour="#2299FF")+
  labs(x=expression(paste(PaO[2]:FiO[2], " Ratio")), y="Hospital Mortality (%)", title="Unadjusted Association of Day 1 PF Ratio and Hospital Mortality")+
  coord_cartesian(ylim = c(0,0.45), clip = "off")+
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        axis.text.y = element_text(colour = c("black", "white", "black", "black", "black", "#2299FF", "#2299FF", "#2299FF", "#2299FF")))+
  scale_y_continuous(breaks = c(seq(0, 0.5, by=0.1), x_100$y, x_200$y, x_300$y, x_700$y),
                     labels = scales::number_format(accuracy = 0.01))+
  scale_x_continuous(breaks = c(seq(0,700, by=100), 50, 150))
```

::: panel-tabset
## Hospital Mortality Curve

```{r, warning=FALSE, message=FALSE}
mort_pf_curve
```

## Hospital Mortality Curve with Cutpoints

```{r, warning=FALSE, message=FALSE}
mort_pf_values
```
:::

```{r, include=FALSE}
# Create Function to determine discrimination characteristics for each outcome
discrim <- function(outcome){
  cp_optimal <- cutpointr(ahrf_validation, x = pf_cont, class = {{outcome}}, pos_class = 1, metric = youden, na.rm = T, method = maximize_metric, boot_runs=10) |> add_metric(list(ppv, npv, youden))

# Bootstrap for AUC 95% CIs
auc_boot <- boot_ci(cp_optimal, AUC)

# Create dataframe of results
pf_cat_test <- cp_optimal |> 
  select(optimal_cutpoint, sensitivity, specificity, ppv, npv, acc, youden...4, AUC) |> 
  mutate_if(is.numeric, round, 3) |> 
  pivot_longer(names_to = "Criteria",
               cols = everything()) |> 
  mutate(Criteria = case_when(
    Criteria == "optimal_cutpoint" ~ "Optimal Cutpoint",
    Criteria == "sensitivity" ~ "Sensitivity (Sn)",
    Criteria == "specificity" ~ "Specificity (Sp)",
    Criteria == "ppv" ~ "Positive Predictive Value (PPV)",
    Criteria == "npv" ~ "Negative Predictive Value (NPV)",
    Criteria == "acc" ~ "Accuracy",
    Criteria == "youden...4" ~ "Youden's Index",
    Criteria == "AUC" ~ "Area Under Curve (AUC)",
  ))

return(pf_cat_test)  
}

# Function for ROC Curve
roc_curve <- function(outcome){
  cp_optimal <- cutpointr(ahrf_validation, x = pf_cont, class = {{outcome}}, pos_class = 1, metric = youden, na.rm = T, method = maximize_metric, boot_runs=10) |>   add_metric(list(ppv, npv, youden))

# Bootstrap for AUC 95% CIs
auc_boot <- boot_ci(cp_optimal, AUC)

# Plot
roc_plot_optimal <- plot_roc(cp_optimal) + 
  theme_minimal()+
  geom_segment(aes(x = 0, xend = 1, y = 0, yend = 1), color="grey", linetype="dashed") +
  annotate("text", x = 0.7, y = 0.05, label = paste0(
    "AUC = ", round(cp_optimal$AUC, 3), " (95% CI = ", round(auc_boot$values[1],3), " - ", round(auc_boot$values[2],3), ")"))+
  annotate("text", x=0.25, y=0.70, label= paste0("Optimal Cutpoint\n PF = ", round(cp_optimal$optimal_cutpoint, 0)))
 
return(roc_plot_optimal) 
}


# Outcomes
# Hospital Mortality
discr_hosp_mort <- discrim(DIED_HOSP) |> rename("Hospital Mortality" = value)
plot_hosp_mort <-roc_curve(DIED_HOSP)+ labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~Hospital~Mortality)

# ICU Mortality
discr_icu_mort <- discrim(DIED_ICU) |> rename("ICU Mortality" = value)
plot_icu_mort <-roc_curve(DIED_ICU)+ labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~ICU~Mortality)

# 28-Day Mortality
discr_28d_mort <- discrim(mort_28d) |> rename("28-Day Mortality" = value)
plot_28d_mort <-roc_curve(mort_28d)+ labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~28-Day~Mortality)

# 90-Day Mortality
discr_90d_mort <- discrim(mort_90d) |> rename("90-Day Mortality" = value)
plot_90d_mort <-roc_curve(mort_90d)+ labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~90-Day~Mortality)

# 180-Day Mortality
discr_180d_mort <- discrim(mort_180d) |> rename("180-Day Mortality" = value)
plot_180d_mort <-roc_curve(mort_180d)+ labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~180-Day~Mortality)

# 1-Year Mortality
discr_365d_mort <- discrim(mort_365d) |> rename("1-Year Mortality" = value)
plot_365d_mort <-roc_curve(mort_365d)+ labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~1-Year~Mortality)



# Table of Test Details
tbl_disc <- discr_hosp_mort |> 
  left_join(discr_icu_mort, by = "Criteria") |> 
  left_join(discr_28d_mort, by = "Criteria") |> 
  left_join(discr_90d_mort, by = "Criteria") |> 
  left_join(discr_180d_mort, by = "Criteria") |> 
  left_join(discr_365d_mort, by = "Criteria") |> 
  gt() |> 
  tab_style(style = cell_text(weight = "bold"), locations = cells_column_labels())
```


### Primary & Secondary Outcomes: Discriminator Characteristics

```{r, warning=FALSE, message=FALSE}
tbl_disc
```

### Primary & Secondary Outcomes: Receiver Operator Curves

::: panel-tabset
## Hospital Mortality
```{r, warning=FALSE, message=FALSE}
plot_hosp_mort
```
## ICU Mortality
```{r}
plot_icu_mort
```
## 28-Day Mortality
```{r}
plot_28d_mort
```
## 90-Day Mortality
```{r}
plot_90d_mort
```
## 180-Day Mortality
```{r}
plot_180d_mort
```
## 1-Year Mortality
```{r}
plot_365d_mort
```
:::

### Subgroup Analysis

```{r, include=FALSE}
# Ventilatory Support
roc_ippv <- ahrf_validation |> 
  filter(INV_IND == 1) |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_noippv <- ahrf_validation |> 
  filter(INV_IND == 0) |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

# Multicurve ROC
rocs_ippv <- list()
rocs_ippv[["Intubated"]] <- roc_ippv
rocs_ippv[["Not Intubated"]] <- roc_noippv

multiroc_ippv <-
  ggroc(rocs_ippv, legacy.axes = TRUE) +
  theme_minimal()+
  labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~Hospital~Mortality,
       subtitle = "In Patients Receiving IPPV",
       x="1 - Specificity", y="Sensitivity", color = " ")+
  geom_abline(intercept = 0, slope = 1, color="grey", linetype="dashed")+
  annotate("text", x = 0.7, y = 0.35, label = bquote(AUC[Intubated]==.(round(rocs_ippv$`Intubated`$auc,3))))+
  annotate("text", x = 0.7, y = 0.25, label = bquote(AUC[`Not Intubated`]==.(round(rocs_ippv$`Not Intubated`$auc,3))))

# Admission Diagnosis
roc_cardiac <- ahrf_validation |> 
  filter(ap3diag_min == "Cardiac Surgery") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_trauma_neurosx <- ahrf_validation |> 
  filter(ap3diag_min == "Trauma/Neurosurgery") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_postop <- ahrf_validation |> 
  filter(ap3diag_min == "Post-Operative") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_medical <- ahrf_validation |> 
  filter(ap3diag_min == "Medical") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_sepsis <- ahrf_validation |> 
  filter(ap3diag_min == "Sepsis") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)



# Multicurve ROC
rocs_adm <- list()
rocs_adm[["Cardiac Surgery"]] <- roc_cardiac
rocs_adm[["Trauma/Neurosurgery"]] <- roc_trauma_neurosx
rocs_adm[["Post-Operative"]] <- roc_postop
rocs_adm[["Medical"]] <- roc_medical
rocs_adm[["Sepsis"]] <- roc_sepsis


multiroc_adm <-
  ggroc(rocs_adm, legacy.axes = TRUE) +
  theme_minimal()+
  labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~Hospital~Mortality,
       subtitle = "By Admission Diagnosis",
       x="1 - Specificity", y="Sensitivity", color = " ")+
  geom_abline(intercept = 0, slope = 1, color="grey", linetype="dashed")+
  annotate("text", x = 0.7, y = 0.40, label = bquote(AUC[`Cardiac Surgery`]==.(round(rocs_adm$`Cardiac Surgery`$auc,3))))+
  annotate("text", x = 0.7, y = 0.30, label = bquote(AUC[`Trauma/Neurosurgery`]==.(round(rocs_adm$`Trauma/Neurosurgery`$auc,3))))+
  annotate("text", x = 0.7, y = 0.20, label = bquote(AUC[`Post-Operative`]==.(round(rocs_adm$`Post-Operative`$auc,3))))+
  annotate("text", x = 0.7, y = 0.10, label = bquote(AUC[`Medical`]==.(round(rocs_adm$`Medical`$auc,3))))+
  annotate("text", x = 0.7, y = 0.00, label = bquote(AUC[`Sepsis`]==.(round(rocs_adm$`Sepsis`$auc,3))))

# Age Categories
roc_age_44 <- ahrf_validation |> 
  filter(age_cat == "<44") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_age_45_64 <- ahrf_validation |> 
  filter(age_cat == "45-64") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_age_65_84 <- ahrf_validation |> 
  filter(age_cat == "65-84") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_age_84 <- ahrf_validation |> 
  filter(age_cat == ">84") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

# Multicurve ROC
rocs_age <- list()
rocs_age[["Age < 44"]] <- roc_age_44
rocs_age[["Age 45-64"]] <- roc_age_45_64
rocs_age[["Age 65-84"]] <- roc_age_65_84
rocs_age[["Age > 84"]] <- roc_age_84


multiroc_age <-
  ggroc(rocs_age, legacy.axes = TRUE) +
  theme_minimal()+
  labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~Hospital~Mortality,
       subtitle = "By Age Category",
       x="1 - Specificity", y="Sensitivity", color = " ")+
  geom_abline(intercept = 0, slope = 1, color="grey", linetype="dashed")+
  annotate("text", x = 0.7, y = 0.40, label = bquote(AUC[`Age < 44`]==.(round(rocs_age$`Age < 44`$auc,3))))+
  annotate("text", x = 0.7, y = 0.30, label = bquote(AUC[`Age 45-64`]==.(round(rocs_age$`Age 45-64`$auc,3))))+
  annotate("text", x = 0.7, y = 0.20, label = bquote(AUC[`Age 65-84`]==.(round(rocs_age$`Age 65-84`$auc,3))))+
  annotate("text", x = 0.7, y = 0.10, label = bquote(AUC[`Age > 84`]==.(round(rocs_age$`Age > 84`$auc,3))))

# Sex
roc_sex_m <- ahrf_validation |> 
  filter(SEX == "Male") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_sex_f <- ahrf_validation |> 
  filter(SEX == "Female") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

# Multicurve ROC
rocs_sex <- list()
rocs_sex[["Male"]] <- roc_sex_m
rocs_sex[["Female"]] <- roc_sex_f

multiroc_sex <-
  ggroc(rocs_sex, legacy.axes = TRUE) +
  theme_minimal()+
  labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~Hospital~Mortality,
       subtitle = "By Patient Sex",
       x="1 - Specificity", y="Sensitivity", color = " ")+
  geom_abline(intercept = 0, slope = 1, color="grey", linetype="dashed")+
  annotate("text", x = 0.7, y = 0.35, label = bquote(AUC[Male]==.(round(rocs_sex$`Male`$auc,3))))+
  annotate("text", x = 0.7, y = 0.25, label = bquote(AUC[`Female`]==.(round(rocs_sex$`Female`$auc,3))))

# Frailty
roc_frailty_fit <- ahrf_validation |> 
  filter(FRAILTY == "Fit/Well") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_frailty_mild <- ahrf_validation |> 
  filter(FRAILTY == "Mild") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_frailty_mod <- ahrf_validation |> 
  filter(FRAILTY == "Moderate") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_frailty_sev <- ahrf_validation |> 
  filter(FRAILTY == "Severe") |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

# Multicurve ROC
roc_frailty <- list()
roc_frailty[["Fit/Well"]] <- roc_frailty_fit
roc_frailty[["Mild"]] <- roc_frailty_mild
roc_frailty[["Moderate"]] <- roc_frailty_mod
roc_frailty[["Severe"]] <- roc_frailty_sev


multiroc_frailty <-
  ggroc(roc_frailty, legacy.axes = TRUE) +
  theme_minimal()+
  labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~Hospital~Mortality,
       subtitle = "By Frailty Category",
       x="1 - Specificity", y="Sensitivity", color = " ")+
  geom_abline(intercept = 0, slope = 1, color="grey", linetype="dashed")+
  annotate("text", x = 0.7, y = 0.40, label = bquote(AUC[`Fit/Well`]==.(round(roc_frailty$`Fit/Well`$auc,3))))+
  annotate("text", x = 0.7, y = 0.30, label = bquote(AUC[`Mild`]==.(round(roc_frailty$`Mild`$auc,3))))+
  annotate("text", x = 0.7, y = 0.20, label = bquote(AUC[`Moderate`]==.(round(roc_frailty$`Moderate`$auc,3))))+
  annotate("text", x = 0.7, y = 0.10, label = bquote(AUC[`Severe`]==.(round(roc_frailty$`Severe`$auc,3))))

# Treatment Limitations
roc_trt_lmt_y <- ahrf_validation |> 
  filter(TREAT_LMT == 1) |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

roc_trt_lmt_n <- ahrf_validation |> 
  filter(TREAT_LMT == 2) |> 
  roc(response = DIED_HOSP, predictor = pf_cont)

# Multicurve ROC
rocs_trt_lmt <- list()
rocs_trt_lmt[["Yes"]] <- roc_trt_lmt_y
rocs_trt_lmt[["No"]] <- roc_trt_lmt_n

multiroc_trt_lmt <-
  ggroc(rocs_trt_lmt, legacy.axes = TRUE) +
  theme_minimal()+
  labs(title=ROC~Curve~of~PaO[2]:FiO[2]~Ratio~'in'~Predicting~Hospital~Mortality,
       subtitle = "By Treatment Limitation Status",
       x="1 - Specificity", y="Sensitivity", color = " ")+
  geom_abline(intercept = 0, slope = 1, color="grey", linetype="dashed")+
  annotate("text", x = 0.7, y = 0.35, label = bquote(AUC[Yes]==.(round(rocs_trt_lmt$`Yes`$auc,3))))+
  annotate("text", x = 0.7, y = 0.25, label = bquote(AUC[`No`]==.(round(rocs_trt_lmt$`No`$auc,3))))
```

::: panel-tabset
## Ventilatory Support

```{r, warning=FALSE, message=FALSE}
multiroc_ippv
```

## Admission Diagnosis

```{r, warning=FALSE, message=FALSE}
multiroc_adm
```

## Age Categories

```{r, warning=FALSE, message=FALSE}
multiroc_age
```

## Sex

```{r, warning=FALSE, message=FALSE}
multiroc_sex
```

## Frailty

```{r, warning=FALSE, message=FALSE}
multiroc_frailty
```

## Treatment Limitations

```{r, warning=FALSE, message=FALSE}
multiroc_trt_lmt
```
:::

## References
